---
title: "final_exam_exercise"
author: "Howie Rosen"
date: "March 6, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Get the data

First we read in the file

```{r read_in_tab_sep_file}
gapM <- read.table(file = "output/combined_gapMinder.tsv", header = TRUE, sep = "\t")
```

## Choose countries

Then we choose three countries to examine

```{r choose_countries, echo=FALSE}
canada <- subset(gapM, (gapM$country=="Canada"))
unitstates <- subset(gapM, (gapM$country=="United States"))
mexico <- subset(gapM, (gapM$country=="Mexico"))
```

## Plot GDP

Now we will plot the GDP by year for the three countries we chose
We can see in the plots that GDP goes up for each country by year

```{r plot_GDP}
plot(x=canada$year, y=canada$gdpPercap, xlab="year", ylab="GDP", main="GDP by Year, Canada" )
plot(x=unitstates$year, y=unitstates$gdpPercap, xlab="year", ylab="GDP", main="GDP by Year, United States" )
plot(x=mexico$year, y=mexico$gdpPercap, xlab="year", ylab="GDP", main="GDP by Year, Mexico" )
```

## Get stats

Examine summary statistics about life expectancy for all the continents
Big differences across continents, highest mean in Oceania

```{r grab_continent_summary_data}
agg_gap_continent_mean <- aggregate(lifeExp~continent, data=gapM, FUN=mean)
agg_gap_continent_min <- aggregate(lifeExp~continent, data=gapM, FUN=min)
agg_gap_continent_max <- aggregate(lifeExp~continent, data=gapM, FUN=max)
summ_continents <- cbind(agg_gap_continent_mean, agg_gap_continent_min[ ,2], agg_gap_continent_max[ , 2])
colnames(summ_continents) <- c("continent", "mean", "min", "max")
summ_continents
```

# Histograms

Now we'll make a couple of histograms illustrating the distribution of life expectancies across countries

```{r histograms}
hist(gapM$lifeExp, xlab = "life expectancy", main = "Life Expectancy")
hist(gapM$lifeExp, breaks = 50, xlab = "life expectancy", main = "Life Expectancy")
```





